Differential expression analysis: A pipeline to analyse time-course gene expression data

### 2019-05-07 scripts/00_introduction.Rmd
Better text + colors
ENH Now plotting top DE genes

@@

While there are many methods that have been proposed for discrete aspects of
time course data, the entire workflow for analysis of such data remains
difficult, particularly for long, developmental time series. Most methods
proposed for time course data are concerned with detecting genes that are
changing over time (differential expression analysis), examples being `edge`
[@storey:significance], functional component analysis based models [@wu:more],
time-course permutation tests [@park:statistical], and multiple testing
strategies to combine singe time point differential expression analysis
[@sun:multiple]. However, with long time course datasets, particularly in
developmental systems, a massive number of genes will show *some* change. For
example, in the \xxx system, over \xxx\% of genes are shown to be changing
over time. The task in these settings is often not to detect changes in genes,
but to categorize into biologically interpretable patterns the vast number of
changes discovered.

>>>>

# TODO list

- Add legends to the plots that don't have.
- Currently have two functions "plot_genes", "plot_centroids", that do exactly
  the same thing. Find a better name for both of them
- On the schema plot, replace clustering schemas with something else.

@@

While there are many methods that have been proposed for discrete aspects of
time course data, the entire workflow for analysis of such data remains
difficult, particularly for long, developmental time series. Most methods
proposed for time course data are concerned with detecting genes that are
changing over time (differential expression analysis), examples being `edge`
[@storey:significance], functional component analysis based models [@wu:more],
time-course permutation tests [@park:statistical], and multiple testing
strategies to combine singe time point differential expression analysis
[@sun:multiple]. However, with long time course datasets, particularly in
developmental systems, a massive number of genes will show *some* change. For
example, in the **xxx** system, over **xxx%** of genes are shown to be changing
over time. The task in these settings is often not to detect changes in genes,
but to categorize into biologically interpretable patterns the vast number of
changes discovered.

### 2019-05-07 scripts/02_differential_expression_analysis.Rmd
ENH Now plotting top DE genes

```{r echo=FALSE, results="hide"}
# Reloading necessary data and preliminary results
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
```

@@

```{r echo=FALSE, results="hide"}
# Set data files and options
data("shoemaker2015")
data = shoemaker2015$data
meta = shoemaker2015$meta

@@

pvalue = qvalues[, "M-K"]
lfc_epicon = log_fold_change_epicon[, "M-K"]

@@

>>>>

```{r echo=FALSE, results="hide"}
# Reloading necessary data and preliminary results
library(devtools)
library(limma)
library(splines)
library(stats)
library(moanin)
source("utils.R")
```

@@

```{r echo=FALSE, results="hide"}
# Set data files and options
data("shoemaker2015")
data = shoemaker2015$data
meta = shoemaker2015$meta
meta = droplevels(meta)

@@

pvalue = qvalues[, "M-K"]
lfc_epicon = log_fold_change_epicon[, "M-K"]
names(lfc_epicon) = row.names(log_fold_change_epicon)

@@

Let's quickly visualize the top 10 DE genes ranked by p-value.

```{r fig.height=8, fig.width=8}
top_DE_genes_pval = names(sort(pvalue)[1:12])
moanin::plot_genes(data[top_DE_genes_pval, ], splines_model,
		   colors=ann_colors$Group, smooth=TRUE)
```

Now visualize the top 10 DE genes by log fold change

```{r fig.height=8, fig.width=8}
top_DE_genes_lfc = names(
    sort(abs(lfc_epicon),
	 decreasing=TRUE)[1:12])
moanin::plot_genes(data[top_DE_genes_lfc, ], splines_model,
		   colors=ann_colors$Group, smooth=TRUE)
